---
title: "Linear Models"
output: github_document
date: "2024-11-08"
---

```{r}
library(tidyverse)
library(p8105.datasets)
```

Do some data cleaning

```{r}
data("nyc_airbnb")
nyc_airbnb=
  nyc_airbnb |> 
  mutate(stars= review_scores_location/2) |> 
  rename(
    borough =neighbourhood_group,
    neighborhood= neighbourhood
  ) |> 
  filter(
    borough != "Staten Island"
  ) |> 
  select(price, stars, borough, neighborhood, room_type) |> 
  mutate(
    borough=fct_infreq(borough),
    room_type=fct_infreq(room_type)
  )
```
adding last mutate step used to change ref category in order of freq instead of ABC order 

## Fit some models
start relatively simple

```{r}
fit= lm(price ~ stars, data=nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

broom::tidy(fit)


fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value)

fit |> 
  broom::glance()
```
most often uses broom::tidy() line of code bc will format in form of a df and can select and stuff like that


Try a bit more complex
```{r}
fit= 
  lm(price~stars + borough, data=nyc_airbnb)

fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(
    term= str_replace(term, "borough", "Borough: ")
  ) |> 
  knitr::kable(digits=3)
  
```


## Some diagnostics
(backtrack to some EDA) 
```{r}
nyc_airbnb |> 
  ggplot(aes(x=stars, y=price))+
  geom_point()+
  stat_smooth(method="lm")
```

Most diagnostics use residuals

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x=resid)) +
  geom_histogram()
```
first modelr:: line creates new df with residuals added to table of nyc_airbnb

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x=borough, y=resid)) +
  geom_violin()+
  ylim(-200,500)
```
highly skewed residuals apparent


Residuals against stars

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x=stars, y=resid)) +
  geom_point()
```

Residuals against fitted values
```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x=pred, y=resid)) +
  geom_point()
```

## Hypothesis testing
for single coefficients, look at the table
```{r}
fit |> 
  broom::tidy()

fit_null = lm(price~stars + borough, data=nyc_airbnb)
fit_alt = lm(price~stars + borough + room_type, data=nyc_airbnb)

```

Look at both 
```{r}
fit_null |> 
  broom::tidy()

fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```
when adding more than one coefficient, use ANOVA (works w nested models ONLY)

## Do effects differ across boroughs

first, use a lot of interactions
```{r}
nyc_airbnb |> 
  lm(price ~ stars*borough + room_type*borough, data= _) |> 
  broom::tidy()
```
if want to know if these interactions matter, need to run ANOVA test w these interactions

Could just also fit separate models
```{r}
nyc_airbnb |> 
  filter(borough== "Manhattan") |> 
  lm(price~stars+room_type, data= _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough== "Brooklyn") |> 
  lm(price~stars+room_type, data= _) |> 
  broom::tidy()
```
only looking at one borough 


Get fancy and use list columns

```{r}
nyc_airbnb |> 
  nest(data= -borough) |> 
  mutate(
    model= map(data, \(x) lm(price~stars+room_type, data=x)),
    results= map(model, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```
e.g., in Bronx, going from 4 to 5 stars will cost you $4.45


what about room type across manhattan neighborhoods?
```{r}
nyc_airbnb |> 
  filter(borough=="Manhattan",
         neighborhood=="Chinatown") |> 
  lm(price~ stars + room_type, data=_) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough=="Manhattan",
         neighborhood=="Chelsea") |> 
  lm(price~ stars + room_type, data=_) |> 
  broom::tidy()
```

Let's do this for all neighborhoods!

```{r}
manhattan_neighborhood_fit_df=
nyc_airbnb |> 
  filter(borough=="Manhattan") |> 
  nest(data= -(borough:neighborhood)) |> 
  mutate(
    model= map(data, \(x) lm(price~stars+room_type, data=x)),
    results=map(model, broom::tidy)
  ) |> 
  select(neighborhood, results) |> 
  unnest(results)
```

Look at the effect of room type
```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x=term, y=estimate))+
  geom_boxplot()
```

looking into outlier of shared rooms costing more than whole apt (ref group)...
```{r}
manhattan_neighborhood_fit_df |> 
    filter(str_detect(term, "room_type")) |> 
    filter(estimate>0)

nyc_airbnb |> 
  filter(
    neighborhood=="NoHo",
    room_type=="Shared room")
```



